---
title: "Cell Type Annotation, integrated data" 
author: "Ebba Rosendal"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    self_contained: true
    highlight: tango
    df_print: paged
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    keep_md: yes
    fig_caption: true
  html_notebook:
    self_contained: true
    highlight: tango
    df_print: paged
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, result='hold',fig.width=12,tidy=TRUE)
knitr::opts_knit$set(progress=TRUE,verbose=TRUE)
```
<style>
h1, .h1, h2, .h2, h3, .h3, h4, .h4 { margin-top: 50px }
p.caption {font-size: 0.9em;font-style: italic;color: grey;margin-right: 10%;margin-left: 10%;text-align: justify}
</style>

***
# Load data
--> Loading libraries to be used for manual annotation (code hidden)

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hold'}
suppressMessages(require(Matrix))
suppressPackageStartupMessages({
  library(Seurat)
  library(cowplot)
  library(ggplot2)
  library(magrittr)
  library(dplyr)
  library(rafalib)
})
```

--> Loading integrated data set (no introns, 10% threshold mt genes, doublets only removed using co-expression of Xist+Eif2s3y)

```{r echo=TRUE, message=FALSE, warning=FALSE}

alldata <- readRDS("../DataInt/DataInt_220825_output/QC_220825_DataInt_220825.rds")

DefaultAssay(alldata) <- "RNA"

DefaultAssay(alldata)
```

```{r}

alldata <- NormalizeData(alldata)

alldata <- FindVariableFeatures(alldata, nfeatures = 3000)

all_genes <- rownames(alldata)

alldata <- ScaleData(alldata, features = all_genes)

rm(all_genes, features, alldata.filt)

gc()

```

```{r, fig.height=5, fig.width=12}
plot_grid(ncol = 2,
          DimPlot(alldata, reduction = "umap", group.by = "orig.ident")+NoAxes(),
          DimPlot(alldata, reduction = "umap", group.by = "CCA_snn_res.1", label = T)+NoAxes()
)

plot_grid(ncol = 2,
          FeaturePlot(alldata, reduction = "umap", features = "nFeature_RNA")+NoAxes(),
          FeaturePlot(alldata, reduction = "umap", features = "percent_mito")+NoAxes()
)
```

```{r, echo= FALSE, include=FALSE}

marker_genes <- FindAllMarkers(alldata, log2FC.threshold = 0.25, test.use = "wilcox", min.pct = 0.1, min.diff.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 50)


marker_genes %>%
    group_by(cluster) %>%
    top_n(-25, p_val_adj) -> top25

marker_genes %>%
    group_by(cluster) %>%
    top_n(-8, p_val_adj) -> top8

marker_genes %>%
    group_by(cluster) %>%
    top_n(-3, p_val_adj) -> top3

```


--> Listing top 8 marker genes for each cluster, visually (unique) and as a list (not unique)

```{r, echo=FALSE, fig.height=4.5, fig.width=8}
mypar(2, 6, mar = c(4, 6, 3, 1))
for (i in unique(top8$cluster)) {
    barplot(sort(setNames(top8$avg_log2FC, top8$gene)[top8$cluster == i], F),
        horiz = T, las = 1, main = paste0(i, " vs. rest"), border = "white", yaxs = "i")
    abline(v = c(0, 0.25), lty = c(1, 2))
}

top8
```


```{r, fig.height=5, fig.width=12}
#This list can be adjusted/made longer:

#Glutameric neurons: Sclc17a7
#GABA neurons: Gad1, Gad2
#Microglia: Ctss + Csf1r + Cx3cr1
#Astrocytes: Aqp4 + Rorb + Fgfr3
#Oligo: Mag + Plp1 + Mog
#OPC: Pdgfra + Cspg4 + Vcan
#VLMC: Dcn + Col1a1
#Pericytes: Abcc9 + Pdgfrb + Vtn
#Endothelial: Cdln5 + Flt1 
#ChP: Ttr + Aqp1
#CD8+NK: Nkg7 + Cd8a +GzmB


my_markers = c("Slc17a7","Gad1", "Gad2",  "Ctss", "Csf1r", "Cx3cr1", "Aqp4", "Rorb", "Fgfr3", "Mag", "Mog", "Plp1", "Pdgfra","Cspg4", "Vcan", "Dcn", "Col1a1", "Abcc9", "Pdgfrb", "Vtn", "Cldn5", "Flt1", "Ttr", "Aqp1", "Nkg7", "Cd8a", "Gzmb")

DotPlot(alldata, features = my_markers, group.by = "CCA_snn_res.1", assay = "RNA", c("lightgrey", "red4"))+ coord_flip()
```


```{r, fig.height=25, fig.width=10}
plot_list <- list()
for (i in my_markers) {
    plot_list[[i]] <- FeaturePlot(alldata, reduction = "umap", features = i,
        ncol = 1, order = T) + NoAxes() + NoGrid()
}

plot_grid(ncol = 3, plotlist = plot_list)
getwd()
```

```{r, fig.height=3, fig.width=12}
VlnPlot(alldata, group.by = "CCA_snn_res.1", features = "nFeature_RNA", pt.size = 0)
VlnPlot(alldata, group.by = "CCA_snn_res.1", features = "percent_mito", pt.size = 0)
VlnPlot(alldata, group.by = "CCA_snn_res.1", features = "percent_ribo", pt.size = 0)
```



```{r, echo=TRUE}

#checking active ident before annotation
levels(alldata)


alldata_annotated <- RenameIdents(alldata, 
                                  '20' = 'Low_quality',
                                  '3' = 'Low_quality',
                                  '31' = 'Doublet',
                                  '2' = 'Neuron_glu',
                                  '4' = 'Neuron_glu',
                                  '5' = 'Neuron_glu',
                                  '8' = 'Neuron_glu',
                                  '9' = 'Neuron_glu',
                                  '10' = 'Neuron_glu',
                                  '13' = 'Neuron_glu',
                                  '14' = 'Neuron_glu',
                                  '16' = 'Neuron_glu',
                                  '19' = 'Neuron_glu',
                                  '22' = 'Neuron_glu',
                                  '24' = 'Neuron_glu',
                                  '26' = 'Neuron_glu',
                                  '6' = 'Neuron_gaba',
                                  '7' = 'Neuron_gaba',
                                  '15' = 'Neuron_gaba',
                                  '18' = 'Neuron_gaba',
                                  '11' = 'Micro',
                                  '12' = 'Micro',
                                  '25' = 'Micro',
                                  '1' = 'Astro',
                                  '0' = 'Oligo',
                                  '17' = 'OPC',
                                  '27' = 'VLMCs',
                                  '23' = 'VLMCs',
                                  '28' = 'Peri',
                                  '21' = 'Endo',
                                  '30' = 'ChP',
                                  '29' = 'T-cells')

#checking active ident after annotation
levels(alldata_annotated)

#Extracting cell type prediction
celltype <- alldata_annotated@active.ident

#Adding new metadata under the name "celltype" 
alldata_annotated <- AddMetaData(alldata_annotated, celltype, col.name='celltype')

#There is probably a nicer way to do this, but it works
```

# Subclustering micro and astro to identify poss. doublets

### Astro

```{r, echo=FALSE}
Astro <- subset(alldata_annotated, subset = celltype == "Astro")

Astro <- FindVariableFeatures(Astro, nfeatures = 2000)
Astro <- ScaleData(Astro)
Astro <- RunPCA(Astro, npcs = 50)
Astro <- RunUMAP(Astro,
                   reduction = "pca", 
                   dims = 1:20,
                   n.components=2,
                   n.neighbors=10,
                   n.epochs=200,
                   min.dist=0.3,
                   learning.rate=1,
                   spread=1)

Astro <- FindNeighbors(Astro, dims = 1:20, k.param = 10)
Astro <- FindClusters(Astro, graph.name = "RNA_snn", resolution = 0.5, algorithm = 1)
```

-> Check various features on UMAP

```{r, fig.height=3, fig.width=10}
plot_grid(ncol = 4, 
          DimPlot(Astro, reduction = "umap", group.by = "ident", label = TRUE) 
          +ggtitle("Clusters")+theme(legend.position='bottom'),
          DimPlot(Astro, reduction = "umap", group.by = "sex", label = FALSE) 
          +ggtitle("Sex")+theme(legend.position='bottom'),
          DimPlot(Astro, reduction = "umap", group.by = "treatment", label = FALSE) 
          +ggtitle("Treatment")+theme(legend.position='bottom'),
          DimPlot(Astro, reduction = "umap", group.by = "genotype", label = FALSE) 
          +ggtitle("Genotype")+theme(legend.position='bottom')
    )

qc <- c("nFeature_RNA", "percent_mito", "percent_ribo", "percent_hb")
plot_list <- list()
for (i in qc) {
    plot_list[[i]] <- FeaturePlot(Astro, reduction = "umap", features = i,
        ncol = 1, order = T) + NoAxes() + NoGrid()
}

plot_grid(ncol = 4, plotlist = plot_list)

```

```{r}
marker_genes <- FindAllMarkers(Astro, log2FC.threshold = 0.25, test.use = "wilcox", 
                               only.pos = TRUE, max.cells.per.ident = 20)
marker_genes <- marker_genes[ which(marker_genes$p_val_adj < 0.05), ]

marker_genes %>%
    group_by(cluster) %>%
    top_n(-8, p_val_adj) -> top8

top8_list <- top8$gene
DoHeatmap(Astro, features = top8_list, group.by = "RNA_snn_res.0.5")
top8
```

```{r, fig.height=4, fig.width=14}
plot_grid(ncol = 4,
          VlnPlot(Astro, group.by = "RNA_snn_res.0.5", features = "nFeature_RNA"),
          VlnPlot(Astro, group.by = "RNA_snn_res.0.5", features = "percent_mito"),
          VlnPlot(Astro, group.by = "RNA_snn_res.0.5", features = "Gfap", assay = "RNA"), 
          VlnPlot(Astro, group.by = "RNA_snn_res.0.5", features = "Slc17a7", assay = "RNA")
)
```

```{r}

Astro.doublet <- Astro[,Astro@active.ident == 6]

rm(Astro)
gc()
```

### Micro


```{r}
Micro <- subset(alldata_annotated, subset = celltype == "Micro")

Micro <- FindVariableFeatures(Micro, nfeatures = 2000, assay = "RNA")
Micro <- ScaleData(Micro, assay = "RNA")
Micro <- RunPCA(Micro, npcs = 50, assay = "RNA")
Micro <- RunUMAP(Micro,
                   reduction = "pca", 
                   dims = 1:20,
                   n.components=2,
                   n.neighbors=10,
                   n.epochs=200,
                   min.dist=0.3,
                   learning.rate=1,
                   spread=1)

Micro <- FindNeighbors(Micro, dims = 1:10, assay = "RNA", k.param =10)
Micro <- FindClusters(Micro, graph.name = "RNA_snn", resolution = 0.5, algorithm = 1)
```

-> Check various features on UMAP

```{r, fig.height=6, fig.width=14}
plot_grid(ncol = 4, 
          DimPlot(Micro, reduction = "umap", group.by = "ident", label = TRUE) 
          +ggtitle("Clusters")+theme(legend.position='bottom'),
          DimPlot(Micro, reduction = "umap", group.by = "sex", label = FALSE) 
          +ggtitle("Sex")+theme(legend.position='bottom'),
          DimPlot(Micro, reduction = "umap", group.by = "treatment", label = FALSE) 
          +ggtitle("Treatment")+theme(legend.position='bottom'),
          DimPlot(Micro, reduction = "umap", group.by = "genotype", label = FALSE) 
          +ggtitle("Genotype")+theme(legend.position='bottom')
    )

qc <- c("nFeature_RNA", "percent_mito", "percent_ribo", "percent_hb")
plot_list <- list()
for (i in qc) {
    plot_list[[i]] <- FeaturePlot(Micro, reduction = "umap", features = i,
        ncol = 1, order = T) + NoAxes() + NoGrid()
}

plot_grid(ncol = 4, plotlist = plot_list)

```

```{r}
marker_genes <- FindAllMarkers(Micro, log2FC.threshold = 0.25, test.use = "wilcox", 
                               only.pos = TRUE, max.cells.per.ident = 20, assay = "RNA")
marker_genes <- marker_genes[ which(marker_genes$p_val_adj < 0.05), ]

marker_genes %>%
    group_by(cluster) %>%
    top_n(-8, p_val_adj) -> top8

top8_list <- top8$gene
DoHeatmap(Micro, features = top8_list, group.by = "RNA_snn_res.0.5")
top8
```

```{r, fig.height=6, fig.width=14}
plot_grid(ncol = 3,
          VlnPlot(Micro, group.by = "RNA_snn_res.0.5", features = "nFeature_RNA"),
          VlnPlot(Micro, group.by = "RNA_snn_res.0.5", features = "percent_ribo"),
          VlnPlot(Micro, group.by = "RNA_snn_res.0.5", features = "percent_mito"),
          VlnPlot(Micro, group.by = "RNA_snn_res.0.5", features = "Ctss", assay = "RNA"),
          VlnPlot(Micro, group.by = "RNA_snn_res.0.5", features = "Mrc1", assay = "RNA"), 
          VlnPlot(Micro, group.by = "RNA_snn_res.0.5", features = "Slc17a7", assay = "RNA")
)
```

```{r}
#No clear cluster with doublets identified

rm(Micro)
gc()
```

***

# Removing low quality cells and doublets

```{r}
# Select cells with low quality and suspected doublets
cluster.remove <- subset(alldata_annotated, subset = celltype == "Low_quality" | celltype == "Doublet")

dim(alldata_annotated)

# Remove selected cells
alldata.clean <- alldata_annotated[, setdiff(WhichCells(alldata_annotated), WhichCells(cluster.remove))]

# Remove potential doublets found by subclustering of micro and astro
alldata.clean <- alldata.clean[, setdiff(WhichCells(alldata.clean),
                                     WhichCells(Astro.doublet))]

dim(alldata.clean)
```

```{r}
rm(alldata, Micro.doublet, astro.dub, alldata_annotated, cluster.remove, marker_genes, plot_list, top25, top3, top8, celltype, qc, top3_list, top8_list)
gc()
```


***

### **Calculating median genes and no. of nuclei per data set**

```{r}
dim(alldata.clean)
median(alldata.clean@meta.data$nFeature_RNA)

wt <- subset(alldata.clean, subset = genotype == "wt")
dim(wt)
median(wt@meta.data$nFeature_RNA)
rm(wt)
gc()

ifnar <- subset(alldata.clean, subset = genotype == "ifnar")
dim(ifnar)
median(ifnar@meta.data$nFeature_RNA)
rm(ifnar)
gc()

none <- subset(alldata.clean, subset = treatment == "none")
dim(none)
median(none@meta.data$nFeature_RNA)
rm(none)
gc()

mock <- subset(alldata.clean, subset = treatment == "mock")
dim(mock)
median(mock@meta.data$nFeature_RNA)
rm(mock)
gc()

lgtv <- subset(alldata.clean, subset = treatment == "lgtv")
dim(lgtv)
median(lgtv@meta.data$nFeature_RNA)
rm(lgtv)
gc()


```


### **Calculating no. of cells/cell type**

--> Calculating absolut no. of cells and % of each cell type in each data set 

```{r, echo=TRUE, fig.height=4, fig.width=10}

#Long and ugly way of figuring out how many cells per cell type in each data set

# rename to make it easier to write the code
dat <- alldata.clean

All <- table(dat$celltype)
wt.no<- table(dat$celltype[dat$genotype == "wt" & dat$inf == "no"])
wt.lgtv <- table(dat$celltype[dat$genotype == "wt" & dat$inf == "yes"])
ifnar.no <- table(dat$celltype[dat$genotype == "ifnar" & dat$inf == "no"])
ifnar.lgtv <- table(dat$celltype[dat$genotype == "ifnar" & dat$inf == "yes"])

# Merging these tables into a matrix for printing
all.cells <- rbind(All, wt.no, wt.lgtv, ifnar.no, ifnar.lgtv)
knitr::kable(all.cells, caption = 'No of cells per celltype')

# Calculating % per celltype per dataset
All <- All *100 /sum(All)
wt.no <- wt.no *100 / sum(wt.no)
wt.lgtv <- wt.lgtv *100 / sum(wt.lgtv)
ifnar.no <- ifnar.no *100 / sum(ifnar.no)
ifnar.lgtv <- ifnar.lgtv *100 / sum(ifnar.lgtv)

# Merging these tables into a matrix for printing
all.perc <- rbind(All, wt.no,  wt.lgtv, ifnar.no, ifnar.lgtv)
knitr::kable(all.perc, caption = '% of cells per celltype', digits = 1)

rm(dat)
gc()

write.csv2(all.perc, file = "Percentage_celltypes_22825.csv")
write.csv2(all.cells, file = "Total_no_celltypes_220825.csv")
```

```{r}
rm(All, all.cells, all.perc, dat, marker_genes, top8_list, i, ifnar.lgtv, ifnar.mock, ifnar.none, wt.lgtv, wt.mock, wt.none)
gc()
```

### **Top 25 genes per annotated cell type**

```{r}
marker_genes <- FindAllMarkers(alldata.clean, log2FC.threshold = 0.25, test.use = "wilcox", min.pct = 0.1, min.diff.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 50)

marker_genes %>%
    group_by(cluster) %>%
    top_n(-25, p_val_adj) -> top25

write.csv2(top25, file = "Top25_markergenes_per_cell_type_220825.csv")
```

### **Write document for manually making DotPlot of selected marker genes in graphpad**

```{r, fig.height=3.5, fig.width=9.5}
# Save dotplot as a file to export values to excel/graphpad for visualization
celltype_markers <- DotPlot(alldata.clean, features = my_markers, group.by = "celltype", assay = "RNA", c("lightgrey", "Red4")) 

# Make DotPlot in R to validate 
DotPlot(alldata.clean, features = my_markers, group.by = "celltype", assay = "RNA", c("lightgrey", "Red3"), scale = TRUE) +theme(axis.text.x = element_text(angle = 45, hjust=1))

write.csv2(celltype_markers$data, file="DotPlot_celltype_markers_220825.csv")

getwd()
```

### **Visualization of annotation on UMAP**

```{r, fig.height=3, fig.width=4}

pdf("UMAP_cell_annotation_220825.pdf", onefile = TRUE, width=5, height=4)

plot_grid(ncol = 1, 
          DimPlot(alldata.clean, reduction = "umap", group.by = "celltype",
                  label=FALSE, cols = c("Red4", "red3", "royalblue4", "royalblue3", "steelblue1", "lightslateblue", "darkgreen", "forestgreen", "palegreen3", "lightgreen", "grey70")) +theme(legend.position='right')
)
dev.off()
getwd()


```

***

# Saving data

```{r save, echo=TRUE, message='hide',warning='hide'}

#dir.create('./ClustAnnotation_220825_output', showWarnings = F)
saveRDS(alldata.clean,"./ClustAnnotation_220825_output/QC_220825_DataInt_220825_Anno_220825.rds")


getwd()
```


***
# Session Info

```{r}
sessionInfo()
```
