---
title: "Data Integration"
author: "Ebba Rosendal"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    self_contained: true
    highlight: tango
    df_print: paged
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    keep_md: yes
    fig_caption: true
  html_notebook:
    self_contained: true
    highlight: tango
    df_print: paged
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, result='hold',fig.width=12,tidy=TRUE)
knitr::opts_knit$set(progress=TRUE,verbose=TRUE)
```
<style>
h1, .h1, h2, .h2, h3, .h3, h4, .h4 { margin-top: 50px }
p.caption {font-size: 0.9em;font-style: italic;color: grey;margin-right: 10%;margin-left: 10%;text-align: justify}
</style>

***
# Load data
--> Loading libraries to be used for data int

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hold'}
suppressPackageStartupMessages({
  library(Seurat)
  library(cowplot)
  library(ggplot2)
  library(magrittr)
  library(dplyr)
  library(rafalib)
})

```

--> Loading merged data set from QC 

```{r, message=FALSE, warning=FALSE}

setwd("C:/Users/ebro0001/OneDrive - Umeå universitet/Skrivbordet/220518_snRNA/DataInt")
alldata <- readRDS("../QC/QC_220825_output/QC_220825.rds")

```

***
# Data integration

--> Splitting the combined object into a list, with each dataset as an object using `SplitObject`. The objects are then individually normalized using `NormalizeData` and highly variable genes detected using `FindVariableFeatures` with `nFeatures=2000`

```{r, message=TRUE, warning=FALSE}
#Splitting merged data set into a list
alldata.list <- SplitObject(alldata, split.by = "orig.ident")

rm(alldata)
gc()

#General pre-processing on objects in list.
#These steps are taken from the example under ´FindIntegrationAnchors´
for (i in 1:length(alldata.list)) {
    alldata.list[[i]] <- NormalizeData(alldata.list[[i]], verbose = FALSE)
    alldata.list[[i]] <- FindVariableFeatures(alldata.list[[i]], 
                                              selection.method = "vst",
                                              nfeatures = 2000,verbose = FALSE)
}

```


--> Identifying anchors using `FindIntegrationAnchors` with `dims = 1:20`and `reduction = "cca"` 

```{r, echo=TRUE}

#Anchors identified using the FindIntegrationAnchors function
alldata.anchors <- FindIntegrationAnchors(object.list = alldata.list, dims = 1:20, reduction = "cca")

```

--> Deleting old file (alldata.list) to clear up memory (code hidden).
 
```{r, echo=FALSE, include=FALSE}
# remove all objects that will not be used.
rm(alldata.list)

# run garbage collect to free up memory
gc()
```

--> Integrate data using `IntegrateData` with anchors identified previously and `dims = 1:20`

Output (gene expression matrix for integrated data) is saved under a new assay name `new.assay.name = "CCA"` 

```{r, echo=TRUE}

#Anchors passed to the IntegrateData function, which returns a Seurat object.
alldata.int <- IntegrateData(anchorset = alldata.anchors, dims = 1:20, new.assay.name = "CCA", verbose = FALSE)

#Check that new assay slot is created under the name `CCA`.
names(alldata.int@assays)

```

--> Deleting old files (alldata.anchors) to clear up memory (code hidden).
 
```{r, echo=FALSE, include=FALSE}
# remove all objects that will not be used.
rm(alldata.anchors)

# run garbage collect to free up memory
gc()
```


### **Run dimensionality reduction on integrated space**

```{r echo=TRUE, message=FALSE, warning=FALSE, results='hold'}

DefaultAssay(alldata.int) <- "CCA"

#Running Dimensionality reduction on integrated space

alldata.int <- ScaleData(alldata.int, verbose = FALSE)
alldata.int <- RunPCA(alldata.int, npcs = 30, verbose = FALSE)
alldata.int <- RunUMAP(alldata.int, 
                   reduction = "pca", 
                   dims = 1:30,
                   n.components=2,
                   n.neighbors=30,
                   n.epochs=200,
                   min.dist=0.3,
                   learning.rate=1,
                   spread=1)
alldata.int <- FindNeighbors(alldata.int, reduction = "pca", dims = 1:30)
alldata.int <- FindClusters(alldata.int, resolution = 1)
```

### **Comparison of un-integrated and the integrated space UMAP**

```{r, echo=FALSE, message='hide', warning='hide',results='hold',fig.height=3, fig.width=8}
plot_grid(ncol = 2,
          DimPlot(alldata.int, reduction = "umap", group.by = "orig.ident")+NoAxes()
          +ggtitle("Integrated"),
          DimPlot(alldata.int, reduction = "umap", group.by = "CCA_snn_res.1", 
                  label = TRUE)+NoAxes()+ggtitle("Clusters") 
)
```

***
# Save integrated data

```{r,message=FALSE, warning=FALSE}

#dir.create('./DataInt_220825_output', showWarnings = F)
saveRDS(alldata.int,"./DataInt_220825_output/QC_220825_DataInt_220825.rds")
```
***
# Session Info

```{r}
sessionInfo()
```

